### 2025-01-26 index.qmd
do it

---
title: "Jilbert's Distinction Centrality"
execute: 
  eval: true
  echo: true
  output: true
  warning: false
  message: false
format: 
   html:
      code-line-numbers: true
bibliography: networks.bib
---

```{r Loading Functions}
   library(igraph)
   library(ggraph)
   library(here)
   library(networkdata)
   library(ggpubr)
   source("distinction.R")
```

According to @jilbert24, the Distinction Centrality of a node is a function of the status of each focal node and the degree of status independence of its alters.

It is given by the difference between $i$'s own status and the average status of $i$'s neighbor in the node deleted subgraph that excludes $i$:

$$
   D_i = S_i(G)  - \frac{\sum_{j \in N(i)} S_j(G - i)}{N} 
$$

Where $S_i(G)$ is the (normalized to have maximum 1.0) eigenvector centrality of node $i$ computed in the full graph $G$, $N(i)$ is the set of $i$'s neighbors, and $S_j(G -i)$ is the eigenvector centrality of node $j$ in the node deleted subgraph of $G$ that excludes $i$. The Distinction of a node thus increases with its own status and decreases to the extent that others would gain status if that node was not around.

```{r Custom Functions}
#| echo: false
   plot.graph <- function(x, l, vs = 12, ts = 6) { #plotting function
      library(ggraph)
      d <- distinction(x)$d
      nc <- rep("", length(d))
      nc[which(d < -0.05)] <- "red"
      nc[which(d > 0.05)] <- "blue"
      nc[which(d <= 0.05 & d >= -0.05)] <- "purple"
      if (is.null(V(x)$name) == TRUE) {
         n <- 1:vcount(x)
         tc <- "white"
         s <- vs
         }
      if (is.null(V(x)$name) == FALSE) {
         n <- V(x)$name
         tc <- nc
         vs <- 0
         }
      p <- ggraph(x, layout = l)
      p <- p + geom_edge_link(color = "steelblue", edge_width = 0.5)
      p <- p + geom_node_point(aes(x = x, y = y), size = vs, color = nc) 
      p <- p + geom_node_text(aes(label = n), size = ts, color = tc)
      p <- p + theme_graph() 
   return(p)
   }

   tab.cent <- function(x, d = 3) { #table function
   library(kableExtra)
   t <- kbl(x, 
       digits = d, row.names = FALSE,
       col.names = c("Node", "Distinction", "Status", "Constraint")) %>% 
      kable_styling(bootstrap_options = c("hover", "condensed", "responsive"))
   return(t)
   }
```

```{r}
    g1 <- make_star(7, mode = "undirected") # star graph
    g2 <- g1 + edge(2,3)
    g2 <- g2 + edge(2,4)
    g2 <- g2 + edge(3,4)
    g2 <- g2 + edge(5,6)
    g2 <- g2 + edge(5,7)
    g2 <- g2 + edge(6,7) # structural fold
    g3 <- g1 + edge(2,4)
    g3 <- g3 + edge(3,4)
    g3 <- g3 + edge(5,6)
    g3 <- g3 + edge(5,7) # intra-connected clusters
    g4 <- g3 + edge(2,7) # inter-connected clusters
```

```{r}
#| fig-cap: "A star graph with seven nodes."
#| fig-height: 5
#| fig-width: 5
   plot.graph(g1, "star")
   ggsave(here("Plots", "star.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in the Star Graph."
   tab.cent(distinction(g1))
```

```{r}
#| fig-cap: "A structural fold with seven nodes."
#| fig-height: 5
#| fig-width: 5
   plot.graph(g2, "kk")
   ggsave(here("Plots", "sfold.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in the Structural Fold."
   tab.cent(distinction(g2))
```

```{r}
#| fig-cap: "Intraconnected Clusters."
#| fig-height: 5
#| fig-width: 5
   plot.graph(g3, "star")
   ggsave(here("Plots", "intra-clusters.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in Intraconnected Clusters."
   tab.cent(distinction(g3))
```

```{r}
#| fig-cap: "Interconnected Clusters."
#| fig-height: 5
#| fig-width: 5
   plot.graph(g4, "star")
   ggsave(here("Plots", "inter-clusters.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in Interconnected Clusters."
   tab.cent(distinction(g4))
```

```{r}
#| fig-cap: "Regular Graph."
#| fig-height: 5
#| fig-width: 5
   set.seed(1234)
   g5 <- sample_k_regular(7, 4)
   plot.graph(g5, "circle")
   ggsave(here("Plots", "regular.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in a Regular Graph."
   tab.cent(distinction(g5))
```

```{r}
#| fig-cap: "Small World Graph."
#| fig-height: 5
#| fig-width: 5
   set.seed(1234)
   g6 <- sample_smallworld(dim = 1, 7, 2, 0.05)
   plot.graph(g6, "circle")
   ggsave(here("Plots", "sw.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in a Small World Graph."
   tab.cent(distinction(g6))
```

```{r}
#| fig-cap: "Karate Club Graph."
#| fig-height: 8
#| fig-width: 8
   set.seed(1234)
   plot.graph(karate, "dh", vs = 8, ts = 4)
   ggsave(here("Plots", "karate.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in the Karate Club."
   tab.cent(distinction(karate))
```

```{r}
#| fig-cap: "Florentine Families Graph."
#| fig-height: 12
#| fig-width: 12
   set.seed(1234)
   iso <- which(degree(flo_marriage) == 0)
   g <- delete_vertices(flo_marriage, iso)
   plot.graph(g, "star", ts = 8)
   ggsave(here("Plots", "flo-marriage.png"), width = 12, height = 12)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in the Florentine Families."
   tab.cent(distinction(g))
```

```{r}
   p <- seq(0.1, 0.9, 0.01)
   r <- rep(0, length(p))
   for (i in 1:length(p)) {
      set.seed(123)
      g <- sample_smallworld(dim = 1, 15, 3, p[i]) 
      d <- distinction(g)$u
      b <- betweenness(g)
      r[i] <- cor(d, b)
      }
   dat <- data.frame(p, r)
   ggscatter(data = dat, "p", "r", add = "loess") 
```

Jilbert's Distinction Equivalence