### 2025-03-01 old.qmd

---
title: "Untitled"
format: html
editor: visual
---

```{r}
#| fig-cap: "A star graph with seven nodes."
#| fig-height: 5
#| fig-width: 5
   plot.graph(g1, "star")
   ggsave(here("Plots", "star.png"), width = 6, height = 6)
```

```{r}
#| fig-cap: "A star graph with seven nodes."
#| fig-height: 5
#| fig-width: 5
   plot.graph.norm(g1, "star")
   ggsave(here("Plots", "star.png"), width = 6, height = 6)
```

```{r}
#| fig-cap: "A structural fold with seven nodes."
#| fig-height: 5
#| fig-width: 5
   plot.graph(g2, "kk")
   ggsave(here("Plots", "sfold.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in the Structural Fold."
   tab.cent(distinction(g2))
```

```{r}
#| fig-cap: "Intraconnected Clusters."
#| fig-height: 5
#| fig-width: 5
   plot.graph(g3, "star")
   ggsave(here("Plots", "intra-clusters.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in Intraconnected Clusters."
   tab.cent(distinction(g3))
```

```{r}
#| fig-cap: "Interconnected Clusters."
#| fig-height: 5
#| fig-width: 5
   plot.graph(g4, "star")
   ggsave(here("Plots", "inter-clusters.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in Interconnected Clusters."
   tab.cent(distinction(g4))
```

```{r}
#| fig-cap: "Regular Graph."
#| fig-height: 5
#| fig-width: 5
   set.seed(1234)
   g5 <- sample_k_regular(7, 4)
   plot.graph(g5, "circle")
   ggsave(here("Plots", "regular.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in a Regular Graph."
   tab.cent(distinction(g5))
```

```{r}
#| fig-cap: "Small World Graph."
#| fig-height: 5
#| fig-width: 5
   set.seed(1234)
   g6 <- sample_smallworld(dim = 1, 7, 2, 0.05)
   plot.graph(g6, "circle")
   ggsave(here("Plots", "sw.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in a Small World Graph."
   tab.cent(distinction(g6))
```

```{r}
#| fig-cap: "Karate Club Graph."
#| fig-height: 8
#| fig-width: 8
   set.seed(1234)
   plot.graph(karate, "dh", vs = 8, ts = 4)
   ggsave(here("Plots", "karate.png"), width = 6, height = 6)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in the Karate Club."
   tab.cent(distinction(karate))
```

```{r}
#| fig-cap: "Florentine Families Graph."
#| fig-height: 12
#| fig-width: 12
   set.seed(1234)
   iso <- which(degree(flo_marriage) == 0)
   g <- delete_vertices(flo_marriage, iso)
   plot.graph(g, "star", ts = 8)
   ggsave(here("Plots", "flo-marriage.png"), width = 12, height = 12)
```

```{r}
#| tbl-cap: "Distinction Centrality Scores of Nodes in the Florentine Families."
   tab.cent(distinction(g))
```

```{r}
   p <- seq(0.1, 0.9, 0.01)
   r <- rep(0, length(p))
   for (i in 1:length(p)) {
      set.seed(123)
      g <- sample_smallworld(dim = 1, 15, 3, p[i]) 
      d <- distinction(g)$u
      b <- betweenness(g)
      r[i] <- cor(d, b)
      }
   dat <- data.frame(p, r)
   ggscatter(data = dat, "p", "r", add = "loess") 
```

Jilbert's Distinction Equivalence