### 2025-03-27 jilbert.qmd

---
title: "Untitled"
format: html
---

```{r}
   library(tidyverse)
   library(dplyr)
   library(tidygraph)
   library(ggraph)
   library(igraph)
```

##Parameter Test Networks

#startest.csv
startest <- read.csv(file.choose(), header=TRUE)
stardf<-as.matrix(startest)
stargraph<-graph_from_edgelist(stardf, directed=FALSE)
distinction(stargraph)
plot(stargraph)

#startest2.csv
star2 <- read.csv(file.choose(), header=TRUE)
star2df<-as.matrix(star2)
stargraph2<-graph_from_edgelist(star2df, directed=FALSE)
distinction(stargraph2)
plot(stargraph2)

#intraclustertest.csv
intracluster <- read.csv(file.choose(), header=TRUE)
intraclusterdf<-as.matrix(intracluster)
intraclustergraph<-graph_from_edgelist(intraclusterdf, directed=FALSE)
distinction(intraclustergraph)
plot(intraclustergraph)

#intraclustertest2.csv
intracluster2 <- read.csv(file.choose(), header=TRUE)
intracluster2df<-as.matrix(intracluster2)
intraclustergraph2<-graph_from_edgelist(intracluster2df, directed=FALSE)
distinction(intraclustergraph2)
plot(intraclustergraph2)

##Medici Network

#MediciEdgeList.csv
medici <- read.csv(file.choose(), header=TRUE)

#AlbizziEditMediciEdgeList.csv
nomedici <- read.csv(file.choose(), header=TRUE)

medicidf<-as.matrix(medici)
medicigraph<-graph_from_edgelist(medicidf, directed=FALSE)

nomedicidf<-as.matrix(nomedici)
nomedicigraph<-graph_from_edgelist(nomedicidf, directed=FALSE)

betweenness(medicigraph)
betweenness(nomedeigenicigraph)

eigen_centrality(medicigraph)
eigen_centrality(nomedicigraph)

```